---
title: "Open-visualizations tutorial for repeated measures in R"
author: "Jordy van Langen"
date: '`r format(Sys.time(), "%B %d, %Y")`'
output:
  pdf_document: default
  keep_md: yes
urlcolor: blue
---


## Background

This tutorial is partly inspired by work from [Allen et al. (2019)](https://wellcomeopenresearch.org/articles/4-63) and was initially created to contribute to a GitHub repository called ['open-visualizations'](https://github.com/jorvlan/open-visualizations). The idea behind the ['open-visualizations'](https://github.com/jorvlan/open-visualizations) repository stems from the fact that (open) science - in general - lacks 'fully' transparent and robust visualizations, i.e., figures have always some form of 'hidden-data'. To overcome this issue, I started experimenting with some code in Python and R and after posting a [Tweet](https://twitter.com/jordyvanlangen/status/1228614021076533249), in which I stated that I was working on ['open-visualizations'](https://github.com/jorvlan/open-visualizations), Neuroscientist Micah Allen [replied](https://twitter.com/micahgallen/status/1228615972749266944) and advised me to check out his work on Rainclouds. After performing two of their tutorials, in [R](https://r-project.org/) and [Python](https://python.org) respectively, I thought of a way to combine the raincloud approach with my own work performed in [Python](https://www.python.org/). Shortly after posting another [Tweet](https://twitter.com/jordyvanlangen/status/1231171552684625921 ), which included some figures that I produced in [R](https://r-project.org/), I received a huge amount of encouraging feedback which led me to writing this markdown document. Next to this markdown document, I have created two other tutorials in Python. Both tutorials are available in my repository [open-visualizations](https://github.com/jorvlan/open-visualizations). 

If you have any questions, suggestions for improvement or identify bugs, please open an issue in the GitHub repository [open-visualizations](https://github.com/jorvlan/open-visualizations). 

If you use my repository for your research, please reference it.

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=6, fig.height=5, fig.path='figs_repmes/',
                      echo=TRUE, warning=FALSE, message=FALSE)
```

## R-version check

```{r check R version, echo = TRUE}
R.version$version.string
```

## Package dependencies

Make sure you have the packages that are needed for this tutorial.

- Install **`plyr`** before **`dplyr`**. If you need functions from both **`plyr`** and **`dplyr`**, please load **`plyr`** first, then **`dplyr`**, otherwise error messages might occur (source: R console). **`Rmisc`** also depends on this package.
- Install **`lattice`** since **`Rmisc`** dependends on this package.
- Install **`ggplot2`**, **`dplyr`**, and **`readr`**.
- Install **`rmarkdown`** to convert this .Rmd template into a variety of formats including HTML, MS Word, PDF, and Beamer (Only required if you do not work in Rstudio).
- Install **`Rmisc`** to perform some basic statistical computations (e.g., calculate mean, median, sd, se, ci).
- Install **`gghalves`** from their GitHub repository since it features the newest options as opposed to the version on CRAN. To install from GitHub, the **`devtools`** package needs to be installed first. **`gghalves`** is a **`ggplot2`** extension for easy plotting of half-half geom combinations. Think half boxplot and half jitterplot, or half violinplot and half dotplot.


## Package references   

- **`plyr`** - [Wickham, 2019](https://cran.r-project.org/web/packages/plyr/index.html)
- **`lattice`** - [Sarkar, 2019](https://cran.r-project.org/web/packages/lattice/index.html)
- **`rmarkdown`** - [Allaire & Xie, 2020](https://cran.r-project.org/web/packages/rmarkdown/index.html)
- **`ggplot2`** - [Wickham et al., 2020](https://cran.r-project.org/web/packages/ggplot2/index.html)
- **`dplyr`** - [Wickham et al., 2020](https://cran.r-project.org/web/packages/dplyr/index.html)
- **`readr`** - [Wickham et al., 2020](https://cran.r-project.org/web/packages/readr/index.html)
- **`Rmisc`** - [Hope, 2013](https://cran.r-project.org/web/packages/Rmisc/index.html)
- **`devtools`** - [Wickham & Hester, 2020](https://cran.r-project.org/web/packages/devtools/index.html)
- **`gghalves`** - [Tiedemann, 2020](https://github.com/erocoar/gghalves)  

## Install packages

```{r install packages, include = TRUE, message = FALSE, eval = FALSE}
packages <- c("plyr", "lattice", "ggplot2", "dplyr", "readr", "rmarkdown","Rmisc")

if (length(setdiff(packages, rownames(installed.packages()))) > 0) {
  install.packages(setdiff(packages, rownames(installed.packages())))  
}

if (!require(devtools)) {
    install.packages("devtools")
}
devtools::install_github('erocoar/gghalves')
```

## Load packages

```{r load packages, include = TRUE, echo = TRUE}
library("plyr")
library("lattice")
library("ggplot2")
library("dplyr")
library("readr")
library("rmarkdown")
library("Rmisc")
library("devtools")
library("gghalves")


# width and height variables for saved plots
w = 6
h = 4

# Define limits of y-axis
y_lim_min = 4
y_lim_max = 7.5

# Make the figure folder if it doesn't exist yet
dir.create('../tutorial_R/figs_repmes/', showWarnings = FALSE)

```


## For this tutorial, we make use of the package `gghalves`. 

The main idea behind `gghalves` is that standard `geom's` aggregate data e.g.,`geom_boxplot`, `geom_violin` and `geom_dotplot` who all tend to be an approximation of symmetry. Given that the space to display information is limited, we can make better use of it by cutting the `geom`s in half and displaying additional `geom`s that e.g. give information about the sample size [Tiedemann, 2020](https://github.com/erocoar/gghalves).

## Figure 1

For this example, we make use of the *iris* dataset, which is freely available in R.

We manipulate the dataset by creating two variables (1) `before` and (2) `after`. 
Specifiy the variable `n` and create a dataframe `d` which includes the variables `y`, `x` and `id`.

```{r dataframe settings, include = TRUE, eval = TRUE, echo = TRUE, message = FALSE}
before = iris$Sepal.Length[1:50]
after = iris$Sepal.Length[51:100]
n <- length(before)
d <- data.frame(y = c(before, after),
                x = rep(c(1,2), each=n),
                id = factor(rep(1:n,2)))
```

Let's create a first very basic figure only showing the individual datapoints.

```{r figure 1, include = TRUE, eval = TRUE, echo = TRUE, message = FALSE }
f1 <- ggplot(data=d, aes(y=y)) +
  
   #Add geom_() objects
   geom_point(aes(x=x), color = "magenta", size = 1.5) +
   
   #Define additional settings
   scale_x_continuous(breaks=c(1,2), labels=c("Before", "After"), limits=c(0, 3)) +
   xlab("Condition") + ylab("Value") +
   ggtitle('Figure 1: Repeated measures individual datapoints') +
   theme_classic() +
   coord_cartesian(ylim=c(y_lim_min, y_lim_max))

f1

ggsave('../tutorial_R/figs_repmes/figure1.png', width = w, height = h)
```

## Figure 2

Let's again create a simple figure, but now with the datapoints connected (i.e., intra-individual trends).

```{r figure 2, include = TRUE, eval = TRUE, echo = TRUE, message = FALSE}
f2 <- ggplot(data=d, aes(y=y)) +
  
   #Add geom_() objects
   geom_point(aes(x=x), color = "magenta", size = 1.5) +
   geom_line(aes(x=x, group=id), color = 'lightgray') +
   
   #Define additional settings
   scale_x_continuous(breaks=c(1,2), labels=c("Before", "After"), limits=c(0, 3)) +
   xlab("Condition") + ylab("Value") +
   ggtitle('Figure 2: Repeated measures with connecting lines') +
   theme_classic()+
   coord_cartesian(ylim=c(y_lim_min, y_lim_max))

f2

ggsave('../tutorial_R/figs_repmes/figure2.png', width = w, height = h)

```

## Figure 3

Let's add some jitter to avoid that datapoints overlap.

```{r create jitter, include = TRUE, eval = TRUE, echo = TRUE, message = FALSE}
set.seed(321)
d$xj <- jitter(d$x, amount=.09)
```

Now we create the the figure again, including jitter.

```{r figure 3, include = TRUE, eval = TRUE, echo = TRUE, message = FALSE}
f3 <- ggplot(data=d, aes(y=y)) +
  
   #Add geom_() objects
   geom_point(aes(x=xj), color = "magenta", size = 1.5) +
   geom_line(aes(x=xj, group=id), color = 'lightgray') +
   
   #Define additional settings
   scale_x_continuous(breaks=c(1,2), labels=c("Before", "After"), limits=c(0, 3)) +
   xlab("Condition") + ylab("Value") +
   ggtitle('Figure 3: Repeated measures with jitter and connections') +
   theme_classic()+
   coord_cartesian(ylim=c(y_lim_min, y_lim_max))

f3

ggsave('../tutorial_R/figs_repmes/figure3.png', width = w, height = h)
```

## Figure 4

Produce another figure, but now defined by different colors for `Before` and `After`.
Just for illustrative purposes, `alpha` was used in `geom_line`.

```{r figure 4, include = TRUE, eval = TRUE, echo = TRUE, message = FALSE}
f4 <- ggplot(data=d, aes(y=y)) +
  
   #Add geom_() objects
   geom_point(data = d %>% filter(x=="1"), aes(x=xj), color = 'dodgerblue', size = 1.5,
              alpha = .6) +
   geom_point(data = d %>% filter(x=="2"), aes(x=xj), color = 'darkorange', size = 1.5, 
              alpha = .6) +
   geom_line(aes(x=xj, group=id), color = 'lightgray', alpha = .3) +
   
   #Define additional settings
   scale_x_continuous(breaks=c(1,2), labels=c("Before", "After"), limits=c(0, 3)) +
   xlab("Condition") + ylab("Value") +
   ggtitle('Figure 4: Repeated measures with jittered datapoints and connections') +
   theme_classic()+
   coord_cartesian(ylim=c(y_lim_min, y_lim_max))

f4

ggsave('../tutorial_R/figs_repmes/figure4.png', width = w, height = h)
```


## Figure 5

Let's add box- and violinplots to create `raincloud` like plots.

```{r figure 5, include = TRUE, eval = TRUE, echo = TRUE, message = FALSE}

f5 <- ggplot(data = d, aes(y = y)) +
  
   #Add geom_() objects
   geom_point(data = d %>% filter(x =="1"), aes(x = xj), color = 'dodgerblue', size = 1.5, 
              alpha = .6) +
   geom_point(data = d %>% filter(x =="2"), aes(x = xj), color = 'darkorange', size = 1.5, 
              alpha = .6) +
   geom_line(aes(x = xj, group = id), color = 'lightgray', alpha = .3) +
  
   geom_half_boxplot(
     data = d %>% filter(x=="1"), aes(x=x, y = y), position = position_nudge(x = -.25), 
     side = "r",outlier.shape = NA, center = TRUE, errorbar.draw = FALSE, width = .2, 
     fill = 'dodgerblue') +
   
   geom_half_boxplot(
     data = d %>% filter(x=="2"), aes(x=x, y = y), position = position_nudge(x = .15), 
     side = "r",outlier.shape = NA, center = TRUE, errorbar.draw = FALSE, width = .2, 
     fill = 'darkorange') +
  
   geom_half_violin(
     data = d %>% filter(x=="1"),aes(x = x, y = y), position = position_nudge(x = -.3), 
     side = "l", fill = 'dodgerblue') +
 
   geom_half_violin(
     data = d %>% filter(x=="2"),aes(x = x, y = y), position = position_nudge(x = .3), 
     side = "r", fill = "darkorange") +
  
  
   #Define additional settings
   scale_x_continuous(breaks=c(1,2), labels=c("Before", "After"), limits=c(0, 3)) +
   xlab("Condition") + ylab("Value") +
   ggtitle('Figure 5: Repeated measures with box- and violin plots') +
   theme_classic()+
   coord_cartesian(ylim=c(y_lim_min, y_lim_max))

f5

ggsave('../tutorial_R/figs_repmes/figure5.png', width = w, height = h)

```
## Figure 5b

```{r figure 5b, include = TRUE, eval = TRUE, echo = TRUE, message = FALSE}

f5b <- ggplot(data = d, aes(y = y)) +
  
   #Add geom_() objects
   geom_point(data = d %>% filter(x =="1"), aes(x = xj), color = 'dodgerblue', size = 1.5, 
              alpha = .5) +
   geom_point(data = d %>% filter(x =="2"), aes(x = xj), color = 'darkorange', size = 1.5, 
              alpha = .5) +
   geom_line(aes(x = xj, group = id), color = 'lightgray', alpha = .5) +
  
   geom_half_boxplot(
     data = d %>% filter(x=="1"), aes(x=x, y = y), position = position_nudge(x = -.25), 
     side = "r",outlier.shape = NA, center = TRUE, errorbar.draw = TRUE, width = .1, 
     fill = 'dodgerblue', alpha = .5) +
   
   geom_half_boxplot(
     data = d %>% filter(x=="2"), aes(x=x, y = y), position = position_nudge(x = .15), 
     side = "r",outlier.shape = NA, center = TRUE, errorbar.draw = TRUE, width = .1, 
     fill = 'darkorange', alpha = .5) +
  
   geom_half_violin(
     data = d %>% filter(x=="1"),aes(x = x, y = y), position = position_nudge(x = 1.3), 
     side = "r", fill = 'dodgerblue', alpha = .5, trim = FALSE) +
 
   geom_half_violin(
     data = d %>% filter(x=="2"),aes(x = x, y = y), position = position_nudge(x = .3), 
     side = "r", fill = "darkorange", alpha = .5, trim = FALSE) +
  
  
   #Define additional settings
   scale_x_continuous(breaks=c(1,2), labels=c("Before", "After"), limits=c(0, 3)) +
   xlab("Condition") + ylab("Value") +
   ggtitle('Figure 5b: Repeated measures with box- and violin plots') +
   theme_classic()+
   coord_cartesian(ylim=c(y_lim_min, y_lim_max))

f5b

ggsave('../tutorial_R/figs_repmes/figure5b.png', width = w, height = h)


```

```{r figure 5c, include = TRUE, eval = TRUE, echo = TRUE, message = FALSE}

f5c <- ggplot(data = d, aes(y = y)) +
  
   #Add geom_() objects
   geom_point(data = d %>% filter(x =="1"), aes(x = xj), color = 'dodgerblue', size = 1.5, 
              alpha = .5) +
   geom_point(data = d %>% filter(x =="2"), aes(x = xj), color = 'darkorange', size = 1.5, 
              alpha = .5) +
   geom_line(aes(x = xj, group = id), color = 'lightgray', alpha = .5) +
  
   geom_half_boxplot(
     data = d %>% filter(x=="1"), aes(x=x, y = y), position = position_nudge(x = 1.21), 
     side = "r",outlier.shape = NA, center = TRUE, errorbar.draw = TRUE, width = .1, 
     fill = 'dodgerblue', alpha = .5) +
   
   geom_half_boxplot(
     data = d %>% filter(x=="2"), aes(x=x, y = y), position = position_nudge(x = .15), 
     side = "r",outlier.shape = NA, center = TRUE, errorbar.draw = TRUE, width = .1, 
     fill = 'darkorange', alpha = .5) +
  
   geom_half_violin(
     data = d %>% filter(x=="1"),aes(x = x, y = y), position = position_nudge(x = 1.3), 
     side = "r", fill = 'dodgerblue', alpha = .5, trim = FALSE) +
 
   geom_half_violin(
     data = d %>% filter(x=="2"),aes(x = x, y = y), position = position_nudge(x = .3), 
     side = "r", fill = "darkorange", alpha = .5, trim = FALSE) +
  
  
   #Define additional settings
   scale_x_continuous(breaks=c(1,2), labels=c("Before", "After"), limits=c(0, 3)) +
   xlab("Condition") + ylab("Value") +
   ggtitle('Figure 5c: Repeated measures with box- and violin plots') +
   theme_classic()+
   coord_cartesian(ylim=c(y_lim_min, y_lim_max))

f5c

ggsave('../tutorial_R/figs_repmes/figure5c.png', width = w, height = h)


```


## Note
A potential downside of the current approach is that it uses a lot of filtering and a lot of `geom`s.

## Add descriptive statistics

- Create a dataframe including:
  - Mean
  - Median
  - Standard deviation
  - Standard error
  - Confidence interval (95 %)

```{r calculate descriptive statistics,  include = TRUE, eval = TRUE, echo = TRUE, message = FALSE}
score_mean_1 <- mean(d$y[1:50])
score_mean_2 <- mean(d$y[51:100])
score_median1 <- median(d$y[1:50])
score_median2 <- median(d$y[51:100])
score_sd_1 <- sd(d$y[1:50])
score_sd_2 <- sd(d$y[51:100])
score_se_1 <- score_sd_1/sqrt(n) #-> adjust your n
score_se_2 <- score_sd_2/sqrt(n) #-> adjust your n
score_ci_1 <- CI(d$y[1:50], ci = 0.95)
score_ci_2 <- CI(d$y[51:100], ci = 0.95)

#Create data frame with 2 rows and 7 columns containing the descriptives
group <- c("x", "z")
N <- c(50, 50)
score_mean <- c(score_mean_1, score_mean_2)
score_median <- c(score_median1, score_median2)
sd <- c(score_sd_1, score_sd_2)
se <- c(score_se_1, score_se_2)
ci <- c((score_ci_1[1] - score_ci_1[3]), (score_ci_2[1] - score_ci_2[3]))

#Create the dataframe
summary_df <- data.frame(group, N, score_mean, score_median, sd, se, ci)
```

## Figure 6

Let's add the items calculated in `summary_df` to the figure

```{r figure 6, include = TRUE, eval = TRUE, echo = TRUE, message = FALSE}
f6 <- ggplot(data = d, aes(y = y)) +
  
   #Add geom_() objects
   geom_point(data = d %>% filter(x =="1"), aes(x = xj), color = 'dodgerblue', size = 1.5, 
              alpha = .6) +
   geom_point(data = d %>% filter(x =="2"), aes(x = xj), color = 'darkorange', size = 1.5, 
              alpha = .6) +
   geom_line(aes(x = xj, group = id), color = 'lightgray', alpha = .3) +
  
   geom_half_boxplot(
     data = d %>% filter(x=="1"), aes(x=x, y = y), position = position_nudge(x = -.28), 
     side = "r",outlier.shape = NA, center = TRUE, errorbar.draw = FALSE, width = .2, 
     fill = 'dodgerblue') +
   
   geom_half_boxplot(
     data = d %>% filter(x=="2"), aes(x=x, y = y), position = position_nudge(x = .18), 
     side = "r",outlier.shape = NA, center = TRUE, errorbar.draw = FALSE, width = .2, 
     fill = 'darkorange') +
  
   geom_half_violin(
     data = d %>% filter(x=="1"),aes(x = x, y = y), position = position_nudge(x = -.3), 
     side = "l", fill = 'dodgerblue') +
 
   geom_half_violin(
     data = d %>% filter(x=="2"),aes(x = x, y = y), position = position_nudge(x = .3), 
     side = "r", fill = "darkorange") +
  
   geom_point(data = d %>% filter(x=="1"), aes(x = x, y = score_mean[1]), 
     position = position_nudge(x = -.13), color = "dodgerblue", alpha = .6, size = 1.5) +
  
   geom_errorbar(data = d %>% filter(x=="1"), aes(x = x, y = score_mean[1], 
     ymin = score_mean[1]-ci[1], ymax = score_mean[1]+ci[1]), 
     position = position_nudge(-.13), 
     color = "dodgerblue", width = 0.05, size = 0.4, alpha = .5) + 
  
   geom_point(data = d %>% filter(x=="2"), aes(x = x, y = score_mean[2]), 
    position = position_nudge(x = .13), color = "darkorange", alpha = .6, size = 1.5)+ 
 
   geom_errorbar(data = d %>% filter(x=="2"), aes(x = x, y = score_mean[2], 
    ymin = score_mean[2]-ci[2], 
    ymax = score_mean[2]+ci[2]), position = position_nudge(.13), color = "darkorange", 
    width = 0.05, size = 0.4, alpha = .5) +
  
  
   #Define additional settings
   scale_x_continuous(breaks=c(1,2), labels=c("Before", "After"), limits=c(0, 3)) +
   xlab("Condition") + ylab("Value") +
   ggtitle('Figure 6: Repeated measures with box- and violin plots') +
   theme_classic()+
   coord_cartesian(ylim=c(y_lim_min, y_lim_max))

f6

ggsave('../tutorial_R/figs_repmes/figure6.png', width = w, height = h)

```

## Figure 7

Optionally you could add a line between the two means.
Define the x-coordinates where the line needs to be drawn.
Note these coordinates can be calculated as 1 - position_nudge() for the first variable and 2 + position_nudge() for the second variable, which in our case is 1 - .13 = .87  and 2 + .13 = 2.13.

```{r figure 7,  include = TRUE, eval = TRUE, echo = TRUE, message = FALSE}
x_tick_means <- c(.87, 2.13)

f7 <- ggplot(data = d, aes(y = y)) +
  
   #Add geom_() objects
   geom_point(data = d %>% filter(x =="1"), aes(x = xj), color = 'dodgerblue', size = 1.5, 
              alpha = .6) +
   geom_point(data = d %>% filter(x =="2"), aes(x = xj), color = 'darkorange', size = 1.5, 
              alpha = .6) +
   geom_line(aes(x = xj, group = id), color = 'lightgray', alpha = .3) +
  
   geom_half_boxplot(
     data = d %>% filter(x=="1"), aes(x=x, y = y), position = position_nudge(x = -.28), 
     side = "r",outlier.shape = NA, center = TRUE, errorbar.draw = FALSE, width = .2, 
     fill = 'dodgerblue') +
   
   geom_half_boxplot(
     data = d %>% filter(x=="2"), aes(x=x, y = y), position = position_nudge(x = .18), 
     side = "r",outlier.shape = NA, center = TRUE, errorbar.draw = FALSE, width = .2, 
     fill = 'darkorange') +
  
   geom_half_violin(
     data = d %>% filter(x=="1"),aes(x = x, y = y), position = position_nudge(x = -.3), 
     side = "l", fill = 'dodgerblue') +
 
   geom_half_violin(
     data = d %>% filter(x=="2"),aes(x = x, y = y), position = position_nudge(x = .3), 
     side = "r", fill = "darkorange") +
  
   geom_point(data = d %>% filter(x=="1"), aes(x = x, y = score_mean[1]), 
     position = position_nudge(x = -.13), color = "dodgerblue", alpha = .6, size = 1.5) +
  
   geom_errorbar(data = d %>% filter(x=="1"), aes(x = x, y = score_mean[1], 
     ymin = score_mean[1]-ci[1], ymax = score_mean[1]+ci[1]), 
     position = position_nudge(-.13), 
     color = "dodgerblue", width = 0.05, size = 0.4, alpha = .6) + 
  
   geom_point(data = d %>% filter(x=="2"), aes(x = x, y = score_mean[2]), 
    position = position_nudge(x = .13), color = "darkorange", alpha = .6, size = 1.5)+ 
 
   geom_errorbar(data = d %>% filter(x=="2"), aes(x = x, y = score_mean[2], 
    ymin = score_mean[2]-ci[2], 
    ymax = score_mean[2]+ci[2]), position = position_nudge(.13), color = "darkorange", 
    width = 0.05, size = 0.4, alpha = .6) +
  
  #Add a line connecting the two means
   geom_line(data = summary_df, aes(x = x_tick_means, y = score_mean), color = 'gray', 
    size = 1) +
  
   #Define additional settings
   scale_x_continuous(breaks=c(1,2), labels=c("Before", "After"), limits=c(0, 3)) +
   xlab("Condition") + ylab("Value") +
   ggtitle('Figure 7: Repeated measures with box- and violin plots and means + ci ') +
   theme_classic()+
   coord_cartesian(ylim=c(y_lim_min, y_lim_max))

f7

ggsave('../tutorial_R/figs_repmes/figure7.png', width = w, height = h)

```

## 2 x 2 repeated measures rainclouds in a #butterfly fashion

As a last step, let's create these figures for a 2 x 2 repeated measures study.

Define an additional variable `z` which has two categories `3` and `4` and create a second jittered variable.

```{r create jitter 2, include = TRUE, eval = TRUE, echo = TRUE, message = FALSE}
before = iris$Sepal.Length[1:50] 
after = iris$Sepal.Length[51:100]
n <- length(before) 
d <- data.frame(y = c(before, after),
               x = rep(c(1,2), each=n),
               z = rep(c(3,4), each=n), 
               id = factor(rep(1:n,2))) 

set.seed(321)
d$xj <- jitter(d$x, amount = .09) 
d$xj_2 <- jitter(d$z, amount = .09)
```


```{r figure 8, include = TRUE, eval = TRUE, echo = TRUE, message = FALSE}
f8 <- ggplot(data = d, aes(y = y)) +
  
   #Add geom_() objects
   geom_point(data = d %>% filter(x =="1"), aes(x = xj), color = 'dodgerblue', size = 1.5, 
              alpha = .6) +
   geom_point(data = d %>% filter(x =="2"), aes(x = xj), color = 'darkorange', size = 1.5, 
              alpha = .6) +
   geom_point(data = d %>% filter(z =="3"), aes(x = xj_2), color = 'dodgerblue', size = 1.5, 
              alpha = .6) + 
   geom_point(data = d %>% filter(z =="4"), aes(x = xj_2), color = 'darkorange', size = 1.5, 
              alpha = .6) +
   geom_line(aes(x = xj, group = id), color = 'lightgray', alpha = .3) +
   geom_line(aes(x = xj_2, group = id), color = 'lightgray', alpha = .3) +
  
   geom_half_boxplot(
     data = d %>% filter(x=="1"), aes(x=x, y = y), position = position_nudge(x = -.35), 
     side = "r",outlier.shape = NA, center = TRUE, errorbar.draw = FALSE, width = .2, 
     fill = 'dodgerblue', alpha = .6) +
   
   geom_half_boxplot(
     data = d %>% filter(x=="2"), aes(x=x, y = y), position = position_nudge(x = -1.16), 
     side = "l",outlier.shape = NA, center = TRUE, errorbar.draw = FALSE, width = .2, 
     fill = 'darkorange', alpha = .6) +
  
   geom_half_boxplot(
     data = d %>% filter(z=="3"), aes(x=z, y = y), position = position_nudge(x = 1.3), 
     side = "r",outlier.shape = NA, center = TRUE, errorbar.draw = FALSE, width = .2, 
     fill = 'dodgerblue', alpha = .6) +
   
   geom_half_boxplot(
     data = d %>% filter(z=="4"), aes(x=z, y = y), position = position_nudge(x = .2), 
     side = "r",outlier.shape = NA, center = TRUE, errorbar.draw = FALSE, width = .2, 
     fill = 'darkorange', alpha = .6) +
  
   geom_half_violin(
     data = d %>% filter(x=="1"),aes(x = x, y = y), position = position_nudge(x = -.40), 
     side = "l", fill = 'dodgerblue', alpha = .6) +
 
   geom_half_violin(
     data = d %>% filter(x=="2"),aes(x = x, y = y), position = position_nudge(x = -1.40), 
     side = "l", fill = "darkorange", alpha = .6) +
  
   geom_half_violin(
     data = d %>% filter(z=="3"),aes(x = z, y = y), position = position_nudge(x = 1.45), 
     side = "r", fill = 'dodgerblue', alpha = .6) +
 
   geom_half_violin(
     data = d %>% filter(z=="4"),aes(x = z, y = y), position = position_nudge(x = .45), 
     side = "r", fill = "darkorange", alpha = .6) +
  
   geom_point(data = d %>% filter(x=="1"), aes(x = x, y = score_mean[1]), 
     position = position_nudge(x = -.13), color = "dodgerblue", alpha = .6, size = 1.5) +
  
   geom_errorbar(data = d %>% filter(x=="1"), aes(x = x, y = score_mean[1], 
     ymin = score_mean[1]-ci[1], ymax = score_mean[1]+ci[1]), 
     position = position_nudge(-.13), 
     color = "dodgerblue", width = 0.05, size = 0.4, alpha = .6) + 
  
   geom_point(data = d %>% filter(x=="2"), aes(x = x, y = score_mean[2]), 
    position = position_nudge(x = -1.1), color = "darkorange", alpha = .6, size = 1.5)+ 
 
   geom_errorbar(data = d %>% filter(x=="2"), aes(x = x, y = score_mean[2], 
    ymin = score_mean[2]-ci[2], 
    ymax = score_mean[2]+ci[2]), position = position_nudge(x = -1.1), color = "darkorange", 
    width = 0.05, size = 0.4, alpha = .6) +

   geom_point(data = d %>% filter(z=="3"), aes(x = z, y = score_mean[1]), 
    position = position_nudge(x = 1.15), color = "dodgerblue", alpha = .5) +
  
   geom_errorbar(data = d %>% filter(z=="3"), aes(x = z, y = score_mean[1], 
    ymin = score_mean[1]-ci[1], 
    ymax = score_mean[1]+ci[1]), position = position_nudge(1.15), 
    color = "dodgerblue", width = 0.05, size = 0.4, alpha = .5)+ 

   geom_point(data = d %>% filter(z=="4"), aes(x = z, y = score_mean[2]), 
    position = position_nudge(x = .15), color = "darkorange", alpha = .5)+

   geom_errorbar(data = d %>% filter(z=="4"), aes(x = z, y = score_mean[2], 
    ymin = score_mean[2]-ci[2], ymax = score_mean[2]+ci[2]), position = position_nudge(.15), 
    color = "darkorange", width = 0.05, size = 0.4, alpha = .5)+
  
  
   #Define additional settings
   scale_x_continuous(breaks=c(1,2,3,4), labels=c("Before", "After","Before", "After"), 
                      limits=c(0, 5))+
   xlab("Condition") + ylab("Value") +
   ggtitle('Figure 8: 2 x 2 Repeated measures with box- and violin plots') +
   theme_classic()+
   coord_cartesian(ylim=c(y_lim_min, y_lim_max))

f8

ggsave('../tutorial_R/figs_repmes/figure8.png', width = w, height = h)
```

## Figure 9

Finally, let's add lines that connect the means of each group.

```{r figure 9, include = TRUE, eval = TRUE, echo = TRUE, message = FALSE}
#First we must again define the x-coordinates of the means.
x_tick_means_x <- c(.87, 2.13) #same as above
x_tick_means_z <- c(2.87, 4.13) #just add 2 for each tick

f9 <- ggplot(data = d, aes(y = y)) +
  
   #Add geom_() objects
   geom_point(data = d %>% filter(x =="1"), aes(x = xj), color = 'dodgerblue', size = 1.5, 
              alpha = .6) +
   geom_point(data = d %>% filter(x =="2"), aes(x = xj), color = 'darkorange', size = 1.5, 
              alpha = .6) +
   geom_point(data = d %>% filter(z =="3"), aes(x = xj_2), color = 'dodgerblue', size = 1.5, 
              alpha = .6) + 
   geom_point(data = d %>% filter(z =="4"), aes(x = xj_2), color = 'darkorange', size = 1.5, 
              alpha = .6) +
   geom_line(aes(x = xj, group = id), color = 'lightgray', alpha = .3) +
   geom_line(aes(x = xj_2, group = id), color = 'lightgray', alpha = .3) +
  
   geom_half_boxplot(
     data = d %>% filter(x=="1"), aes(x=x, y = y), position = position_nudge(x = -.35), 
     side = "r",outlier.shape = NA, center = TRUE, errorbar.draw = FALSE, width = .2, 
     fill = 'dodgerblue', alpha = .6) +
   
   geom_half_boxplot(
     data = d %>% filter(x=="2"), aes(x=x, y = y), position = position_nudge(x = -1.16), 
     side = "l",outlier.shape = NA, center = TRUE, errorbar.draw = FALSE, width = .2, 
     fill = 'darkorange', alpha = .6) +
  
   geom_half_boxplot(
     data = d %>% filter(z=="3"), aes(x=z, y = y), position = position_nudge(x = 1.3), 
     side = "r",outlier.shape = NA, center = TRUE, errorbar.draw = FALSE, width = .2, 
     fill = 'dodgerblue', alpha = .6) +
   
   geom_half_boxplot(
     data = d %>% filter(z=="4"), aes(x=z, y = y), position = position_nudge(x = .2), 
     side = "r",outlier.shape = NA, center = TRUE, errorbar.draw = FALSE, width = .2, 
     fill = 'darkorange', alpha = .6) +
  
   geom_half_violin(
     data = d %>% filter(x=="1"),aes(x = x, y = y), position = position_nudge(x = -.40), 
     side = "l", fill = 'dodgerblue', alpha = .6) +
 
   geom_half_violin(
     data = d %>% filter(x=="2"),aes(x = x, y = y), position = position_nudge(x = -1.40), 
     side = "l", fill = "darkorange", alpha = .6) +
  
   geom_half_violin(
     data = d %>% filter(z=="3"),aes(x = z, y = y), position = position_nudge(x = 1.45), 
     side = "r", fill = 'dodgerblue', alpha = .6) +
 
   geom_half_violin(
     data = d %>% filter(z=="4"),aes(x = z, y = y), position = position_nudge(x = .45), 
     side = "r", fill = "darkorange", alpha = .6) +
  
   geom_point(data = d %>% filter(x=="1"), aes(x = x, y = score_mean[1]), 
     position = position_nudge(x = -.13), color = "dodgerblue", alpha = .6, size = 1.5) +
  
   geom_errorbar(data = d %>% filter(x=="1"), aes(x = x, y = score_mean[1], 
     ymin = score_mean[1]-ci[1], ymax = score_mean[1]+ci[1]), 
     position = position_nudge(-.13), 
     color = "dodgerblue", width = 0.05, size = 0.4, alpha = .6) + 
  
   geom_point(data = d %>% filter(x=="2"), aes(x = x, y = score_mean[2]), 
    position = position_nudge(x = .13), color = "darkorange", alpha = .6, size = 1.5)+ 
 
   geom_errorbar(data = d %>% filter(x=="2"), aes(x = x, y = score_mean[2], 
    ymin = score_mean[2]-ci[2], 
    ymax = score_mean[2]+ci[2]), position = position_nudge(x = .13), color = "darkorange", 
    width = 0.05, size = 0.4, alpha = .6) +

   geom_point(data = d %>% filter(z=="3"), aes(x = z, y = score_mean[1]), 
    position = position_nudge(x = -.13), color = "dodgerblue", alpha = .5) +
  
   geom_errorbar(data = d %>% filter(z=="3"), aes(x = z, y = score_mean[1], 
    ymin = score_mean[1]-ci[1], 
    ymax = score_mean[1]+ci[1]), position = position_nudge(-.13), 
    color = "dodgerblue", width = 0.05, size = 0.4, alpha = .5)+ 

   geom_point(data = d %>% filter(z=="4"), aes(x = z, y = score_mean[2]), 
    position = position_nudge(x = .13), color = "darkorange", alpha = .5)+

   geom_errorbar(data = d %>% filter(z=="4"), aes(x = z, y = score_mean[2], 
    ymin = score_mean[2]-ci[2], ymax = score_mean[2]+ci[2]), 
    position = position_nudge(.13), 
    color = "darkorange", width = 0.05, size = 0.4, alpha = .5)+
  
   #Add lines connecting the two means
   geom_line(data = summary_df, aes(x = x_tick_means_x, y = score_mean), 
             color = 'gray', size = 1) +
   geom_line(data = summary_df, aes(x = x_tick_means_z, y = score_mean), 
             color = 'gray', size = 1) +
  
   #Define additional settings
   scale_x_continuous(breaks=c(1,2,3,4), labels=c("Before", "After","Before", "After"), 
                      limits=c(0, 5))+
   xlab("Condition") + ylab("Value") +
   ggtitle('Figure 9: 2 x 2 Repeated measures with box- and violin plots') +
   theme_classic()+
   coord_cartesian(ylim=c(y_lim_min, y_lim_max))

f9

ggsave('../tutorial_R/figs_repmes/figure9.png', width = w, height = h)

```
## Figure 10

## Here, we add an additional variable 'during' if you want to show the effect measured over 3 timepoints.
```{r figure 10, include = TRUE, eval = TRUE, echo = TRUE, message = FALSE}
before = iris$Sepal.Length[1:50]
during = iris$Sepal.Length[51:100]
after = iris$Sepal.Length[1:50]
n <- length(before) 
d <- data.frame(y = c(before, during, after),
               x = rep(c(1,2,3), each=n),
               id = factor(rep(1:n,3)))

set.seed(321)
d$xj <- jitter(d$x, amount = .09) 

f10 <- ggplot(data=d, aes(y=y)) +
  
   #Add geom_() objects
   geom_point(data = d %>% filter(x=="1"), aes(x=xj), color = 'dodgerblue', size = 1.5,
              alpha = .6) +
   geom_point(data = d %>% filter(x=="2"), aes(x=xj), color = 'darkgreen', size = 1.5,
              alpha = .6) +
   geom_point(data = d %>% filter(x=="3"), aes(x=xj), color = 'darkorange', size = 1.5, 
              alpha = .6) +
   
   geom_line(aes(x=xj, group=id), color = 'lightgray', alpha = .3) +
   
   #Define additional settings
   scale_x_continuous(breaks=c(1,2,3), labels=c("Before", "During", "After"), limits=c(0.5, 3.5)) +
   xlab("Condition") + ylab("Value") +
   ggtitle('Figure 10: Repeated measures with jittered datapoints and connections') +
   theme_classic()+
   coord_cartesian(ylim=c(3.5, 7.5))

f10

ggsave('../tutorial_R/figs_repmes/figure10.png', width = w, height = h)



```

```{r figure 11, include = TRUE, eval = TRUE, echo = TRUE, message = FALSE}

before = iris$Sepal.Width[1:50]
during = iris$Sepal.Length[51:100]
after = iris$Sepal.Length[1:50]
n <- length(before) 
d <- data.frame(y = c(before, during, after),
               x = rep(c(1,2,3), each=n),
               id = factor(rep(1:n,3)))

set.seed(321)
d$xj <- jitter(d$x, amount = .09) 


f11 <- ggplot(data=d, aes(y=y)) +
  
   #Add geom_() objects
   geom_point(data = d %>% filter(x=="1"), aes(x=xj), color = 'dodgerblue', size = 1.5,
              alpha = .6) +
   geom_point(data = d %>% filter(x=="2"), aes(x=xj), color = 'darkgreen', size = 1.5,
              alpha = .6) +
   geom_point(data = d %>% filter(x=="3"), aes(x=xj), color = 'darkorange', size = 1.5, 
              alpha = .6) +
   
   geom_line(aes(x=xj, group=id), color = 'lightgray', alpha = .3) +
  
   geom_half_violin(
     data = d %>% filter(x=="1"),aes(x = x, y = y), position = position_nudge(x = 2.2), 
     side = "r", fill = 'dodgerblue', alpha = .5, color = "dodgerblue", trim = TRUE) +
 
   geom_half_violin(
     data = d %>% filter(x=="2"),aes(x = x, y = y), position = position_nudge(x = 1.2), 
     side = "r", fill = "darkgreen", alpha = .5, color = "darkgreen", trim = TRUE) +
   
   geom_half_violin(
     data = d %>% filter(x=="3"),aes(x = x, y = y), position = position_nudge(x = 0.2), 
     side = "r", fill = "darkorange", alpha = .5, color = "darkorange", trim = TRUE) +
   
   #Define additional settings
   scale_x_continuous(breaks=c(1,2,3), labels=c("Before", "During", "After"), limits=c(0.5, 4)) +
   xlab("Condition") + ylab("Value") +
   ggtitle('Figure 11: Repeated measures with jittered datapoints and connections') +
   theme_classic()+
   coord_cartesian(ylim=c(2.3, 7.2))

f11

ggsave('../tutorial_R/figs_repmes/figure11.png', width = w, height = h)
```
```{r figure 12, include = TRUE, eval = TRUE, echo = TRUE, message = FALSE}
f12 <- ggplot(data=d, aes(y=y)) +
  
   #Add geom_() objects
   geom_point(data = d %>% filter(x=="1"), aes(x=xj), color = 'dodgerblue', size = 1.5,
              alpha = .8) +
   geom_point(data = d %>% filter(x=="2"), aes(x=xj), color = 'darkgreen', size = 1.5,
              alpha = .8) +
   geom_point(data = d %>% filter(x=="3"), aes(x=xj), color = 'darkorange', size = 1.5, 
              alpha = .8) +
   
   geom_line(aes(x=xj, group=id), color = 'lightgray', alpha = .3) +
  
   geom_violin(
     data = d %>% filter(x=="1"),aes(x = x, y = y), position = position_nudge(x = 0),
     fill = 'dodgerblue', alpha = .1, color = "white", trim = TRUE) +
 
   geom_violin(
     data = d %>% filter(x=="2"),aes(x = x, y = y), position = position_nudge(x = 0),
     fill = "darkgreen", alpha = .1, color = "white", trim = FALSE) +
   
   geom_violin(
     data = d %>% filter(x=="3"),aes(x = x, y = y), position = position_nudge(x = 0),
     fill = "darkorange", alpha = .1, color = "white", trim = TRUE) +
   
   #Define additional settings
   scale_x_continuous(breaks=c(1,2,3), labels=c("Before", "During", "After"), limits=c(0.5, 4)) +
   xlab("Condition") + ylab("Value") +
   ggtitle('Figure 12: Repeated measures with jittered datapoints and connections') +
   theme_classic()+
   coord_cartesian(ylim=c(2.3, 7.2))

f12

ggsave('../tutorial_R/figs_repmes/figure12.png', width = w, height = h)
```



```{r figure 13, include = TRUE, eval = TRUE, echo = TRUE, message = FALSE}
before = iris$Sepal.Width[1:50]
during = iris$Sepal.Length[51:100]
after = iris$Sepal.Length[1:50]
n <- length(before) 
d <- data.frame(y = c(before, during, after),
               x = rep(c(1,2,3), each=n),
               z = rep(c(4,5,6), each=n), 
               id = factor(rep(1:n,3)))

set.seed(321)
d$xj <- jitter(d$x, amount = .09) 
d$xj_2 <- jitter(d$z, amount = .09)
#d$xj_3 <- jitter(d$a, amount = .09)

f13 <- ggplot(data=d, aes(y=y)) +
  
   #Add geom_() objects
   geom_point(data = d %>% filter(x=="1"), aes(x=xj), color = 'dodgerblue', size = 1.5,
              alpha = .6) +
   geom_point(data = d %>% filter(x=="2"), aes(x=xj), color = 'darkgreen', size = 1.5,
              alpha = .6) +
   geom_point(data = d %>% filter(x=="3"), aes(x=xj), color = 'darkorange', size = 1.5, 
              alpha = .6) +
   
    #Add geom_() objects
   geom_point(data = d %>% filter(z=="4"), aes(x=xj_2), color = 'dodgerblue', size = 1.5,
              alpha = .6) +
   geom_point(data = d %>% filter(z=="5"), aes(x=xj_2), color = 'darkgreen', size = 1.5,
              alpha = .6) +
   geom_point(data = d %>% filter(z=="6"), aes(x=xj_2), color = 'darkorange', size = 1.5, 
              alpha = .6) +
   
   geom_line(aes(x=xj, group=id), color = 'lightgray', alpha = .3) +
   geom_line(aes(x=xj_2, group=id), color = 'lightgray', alpha = .3) +
  
   geom_half_violin(
     data = d %>% filter(x=="1"),aes(x = x, y = y), position = position_nudge(x = -0.2), 
     side = "l", fill = 'dodgerblue', alpha = .5, color = "dodgerblue", trim = TRUE) +
 
   geom_half_violin(
     data = d %>% filter(x=="2"),aes(x = x, y = y), position = position_nudge(x = -1.2), 
     side = "l", fill = "darkgreen", alpha = .5, color = "darkgreen", trim = TRUE) +
   
   geom_half_violin(
     data = d %>% filter(x=="3"),aes(x = x, y = y), position = position_nudge(x = -2.2), 
     side = "l", fill = "darkorange", alpha = .5, color = "darkorange", trim = TRUE) +
   
   geom_half_violin(
     data = d %>% filter(z=="4"),aes(x = x, y = y), position = position_nudge(x = 5.2), 
     side = "r", fill = 'dodgerblue', alpha = .5, color = "dodgerblue", trim = TRUE) +
 
   geom_half_violin(
     data = d %>% filter(z=="5"),aes(x = x, y = y), position = position_nudge(x = 4.2), 
     side = "r", fill = "darkgreen", alpha = .5, color = "darkgreen", trim = TRUE) +
   
   geom_half_violin(
     data = d %>% filter(z=="6"),aes(x = x, y = y), position = position_nudge(x = 3.2), 
     side = "r", fill = "darkorange", alpha = .5, color = "darkorange", trim = TRUE) +
   
   #Define additional settings
   scale_x_continuous(breaks=c(1,2,3,4,5,6), labels=c("Before", "During", "After", "Before", "During", "After"), limits=c(0, 7)) +
   xlab("Condition") + ylab("Value") +
   ggtitle('Figure 13: Repeated measures with jittered datapoints and connections') +
   theme_classic()+
   coord_cartesian(ylim=c(2.3, 7.2))

f13

ggsave('../tutorial_R/figs_repmes/figure13.png', width = w, height = h)
```
```{r figure 14, include = TRUE, eval = TRUE, echo = TRUE, message = FALSE}
f14 <- ggplot(data=d, aes(y=y)) +
  
   #Add geom_() objects
   geom_point(data = d %>% filter(x=="1"), aes(x=xj), color = 'dodgerblue', size = 1.5,
              alpha = .8) +
   geom_point(data = d %>% filter(x=="2"), aes(x=xj), color = 'darkgreen', size = 1.5,
              alpha = .8) +
   geom_point(data = d %>% filter(x=="3"), aes(x=xj), color = 'darkorange', size = 1.5, 
              alpha = .8) +
   
    #Add geom_() objects
   geom_point(data = d %>% filter(z=="4"), aes(x=xj_2), color = 'dodgerblue', size = 1.5,
              alpha = .8) +
   geom_point(data = d %>% filter(z=="5"), aes(x=xj_2), color = 'darkgreen', size = 1.5,
              alpha = .8) +
   geom_point(data = d %>% filter(z=="6"), aes(x=xj_2), color = 'darkorange', size = 1.5, 
              alpha = .8) +
   
   geom_line(aes(x=xj, group=id), color = 'lightgray', alpha = .3) +
   geom_line(aes(x=xj_2, group=id), color = 'lightgray', alpha = .3) +
  
   geom_half_violin(
     data = d %>% filter(x=="1"),aes(x = x, y = y), position = position_nudge(x = 0), 
     side = "r", fill = 'dodgerblue', alpha = .15, color = "lightgray", trim = TRUE) +
 
   geom_half_violin(
     data = d %>% filter(x=="2"),aes(x = x, y = y), position = position_nudge(x = 0), 
     side = "r", fill = "darkgreen", alpha = .15, color = "lightgray", trim = TRUE) +
   
   geom_half_violin(
     data = d %>% filter(x=="3"),aes(x = x, y = y), position = position_nudge(x = 0), 
     side = "r", fill = "darkorange", alpha = .15, color = "lightgray", trim = TRUE) +
   
   geom_half_violin(
     data = d %>% filter(z=="4"),aes(x = x, y = y), position = position_nudge(x = 3), 
     side = "r", fill = 'dodgerblue', alpha = .15, color = "lightgray", trim = TRUE) +
 
   geom_half_violin(
     data = d %>% filter(z=="5"),aes(x = x, y = y), position = position_nudge(x = 3), 
     side = "r", fill = "darkgreen", alpha = .15, color = "lightgray", trim = TRUE) +
   
   geom_half_violin(
     data = d %>% filter(z=="6"),aes(x = x, y = y), position = position_nudge(x = 3), 
     side = "r", fill = "darkorange", alpha = .15, color = "lightgray", trim = TRUE) +
   
   #Define additional settings
   scale_x_continuous(breaks=c(1,2,3,4,5,6), labels=c("Before", "During", "After", "Before", "During", "After"), limits=c(0, 7)) +
   xlab("Condition") + ylab("Value") +
   ggtitle('Figure 14: Repeated measures with jittered datapoints and connections') +
   theme_classic()+
   coord_cartesian(ylim=c(2.3, 7.2))

f14

ggsave('../tutorial_R/figs_repmes/figure14.png', width = w, height = h)
```


## General remarks / tips

- To be more flexible in assigning labels to your figures, the **`ggtext`** package by [Wilke, 2020](https://github.com/wilkelab/ggtext) might be worthwile .
- If you would like to be flexible in plotting multiple figures next to each other, check-out the **`patchwork`** package by [Pedersen, 2020](https://cran.r-project.org/web/packages/patchwork/index.html).
- If you want to save your figures in a high-quality manner (> GB) for e.g., publications, you could save your figure with a `.tiff` extension and add `dpi=` as used in the following line of code:

```{r high-quality figure, include = TRUE, echo = TRUE, eval = FALSE}

ggsave("figure.tiff", height=h, width=w, units='in', dpi=600)

```

- If for some reason your code does not work due to e.g., an update in a package, you could use the following line of code to unload and load that package again.

```{r package reload, include = TRUE, echo = TRUE, eval = FALSE}

if("package_name" %in% (.packages())){
  detach('package:package_name', unload=TRUE)}

library("package_name")

```

You have reached the end of this document.

I hope you'll be able to use this tutorial to create more `open-visualizations` for your research!

If you use this tutorial, please cite it in your work.

[open-visualizations](https://github.com/jorvlan/open-visualizations) for repeated measures in R by [**Jordy van Langen**](https://jordyvanlangen.com) 